Density dependence

### 2020-09-15 R/density dependence.Rmd
added survival analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(lme4)
library(bbmle)
library(car)
library(broom)
library(broom.mixed)
library(patchwork)

@@

```{r}
ha <-
  ha %>% 
  group_by(ha_id_number) %>% 
  arrange(ha_id_number, year) %>% 
  mutate(across(starts_with("density"), lag)) %>% 
  ungroup() %>% 
  #scale year
  mutate(year_post = year - min(year))
```

@@

```{r}
ha_cf <- 
  ha %>% 
  filter(habitat == "CF") %>% 
  mutate(log_ht = log(ht),
         log_ht_prev = log(ht_prev))
```

@@

```{r}
mpois <- glm(shts ~ log(shts_prev) * density_0_10,
             family = poisson(link = "log"),
             data = ha_cf)
lm <- lm(shts ~ log(shts_prev) * density_0_10,
         data = ha_cf)
mnorm <- glm(shts ~ log(shts_prev) * density_0_10,
             family = gaussian(link = "log"),
             data = ha_cf)
mpois2 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|year_post) + (1|plot),
                family = poisson(link = "log"),
                data = ha_cf)

mnorm2 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|year_post) + (1|plot),
                family = gaussian(link = "log"),
                data = ha_cf)

AICtab(mpois, mnorm, lm, mpois2, mnorm2)
```

Lognormal distribution fits best, even after attempting to account of overdispersion with random effects.  I guess despite being discrete, this isn't really "count" data *per se*.

Model doesn't converge though.

```{r}
mnorm3 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|year_post),
               family = gaussian(link = "log"),
               data = ha_cf)

@@

```{r}
mnorm4 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|year_post) + (1|plot),
               family = gaussian(link = "log"),
               data = ha_cf)
AICtab(mnorm2, mnorm4)
```

@@

```{r}
mnorm5 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|plot),
               family = gaussian(link = "log"),
               data = ha_cf)
AICtab(mnorm2, mnorm3, mnorm4, mnorm5)
```

Can't remove year as random effect.

Inspect residuals:

```{r}
ha_aug <- augment(mnorm2)
h <-
  ggplot(ha_aug, aes(.resid)) +
  geom_histogram()

qq <-
  ggplot(ha_aug, aes(sample = .resid)) +
  geom_qq() + geom_qq_line()

resid<- 
  ggplot(ha_aug, aes(x = .fitted, y = .resid)) + geom_point() 
h/(qq|resid)
```
I don't love this, but it is a better fit than a poisson distribution.

@@

- ~~Density should be total shoot number~~
- ~~Should actually be density in previous year~~
- ~~Shoots should be poisson distribution, not log-transformed?~~
- incorporate 1-ha fragments and use habitat type as a predictor.
- Implement FLM approach to density dependence.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(lme4)
library(bbmle)
library(car)
library(broom)
library(broom.mixed)
library(patchwork)
library(performance)

@@

```{r}
ha <-
  ha %>% 
  group_by(ha_id_number) %>% 
  arrange(ha_id_number, year) %>% 
  # mutate(across(starts_with("density"), lag, .names = "prev_{col}")) %>%
  mutate(across(starts_with("density"), lag)) %>% 
  ungroup() %>% 
  # filter(!is.na(prev_density_0_10)) %>% 
  #scale year
  mutate(year_post = year - min(year))
```

@@

```{r}
ha_cf <- 
  ha %>% 
  filter(habitat == "CF") %>% 
  mutate(log_shts = log(shts),
         log_shts_prev = log(shts_prev))

```

@@

## Distribution

Number of shoots is discrete, but not *really* count data.  I'll try a poisson distribution and a lognormal distribution.

```{r}
mpois <- glm(shts ~ log_shts_prev + density_0_10 + density_10_20,
             family = poisson(link = "log"),
             data = ha_cf)
lm <- lm(shts ~ log_shts_prev + density_0_10 + density_10_20,
         data = ha_cf)
mnorm <- glm(shts ~ log_shts_prev + density_0_10 + density_10_20,
             family = gaussian(link = "log"),
             data = ha_cf)
mpois2 <- glmer(shts ~ log_shts_prev + density_0_10 + density_10_20 + (1|ha_id_number) + (1|year_post) + (1|plot),
                family = poisson(link = "log"),
                data = ha_cf)

check_overdispersion(mpois2) #actually underdispersed

mnorm2 <- glmer(shts ~ log_shts_prev + density_0_10 + density_10_20 + (1|ha_id_number) + (1|year_post) + (1|plot),
                family = gaussian(link = "log"),
                data = ha_cf)

AICtab(mpois, mnorm, lm, mpois2, mnorm2)
compare_performance(mpois, mnorm, mpois2, mnorm2, rank = TRUE)
```

Lognormal distribution fits best according to AIC, although poisson has lower RMSE.  I guess despite being discrete, this isn't really "count" data *per se*.  The lognormal model doesn't problems with singularity that the poisson model has.  Both have problems with heteroscedasticity.

```{r}
check_convergence(mnorm2)
check_singularity(mpois2)
check_singularity(mnorm2)
x<-check_heteroscedasticity(mnorm2)
plot(x)
```

## Random effects

```{r}
mnorm3 <- glmer(shts ~ log_shts_prev * density_0_10 + (1|ha_id_number) + (1|year_post),
               family = gaussian(link = "log"),
               data = ha_cf)

@@

```{r}
mnorm4 <- glmer(shts ~ log_shts_prev * density_0_10 + (1|year_post) + (1|plot),
               family = gaussian(link = "log"),
               data = ha_cf)
AICtab(mnorm2, mnorm4)
```

@@

```{r}
mnorm5 <- glmer(shts ~ log_shts_prev * density_0_10 + (1|ha_id_number) + (1|plot),
               family = gaussian(link = "log"),
               data = ha_cf)
AICtab(mnorm2, mnorm3, mnorm4, mnorm5)
```

## Does density matter beyond immediate surroundings?

```{r}
Anova(mnorm2)
```

Apparently not! Weird! It is highly significant when I was accidentally modeling density in the same year. This is not due to removing rows from the first year either.  This is somewhat concerning!

# Survival

```{r}
m_surv <- glmer(surv ~ log_shts_prev + density_0_10 + density_10_20 + (1|plot) + (1|ha_id_number) + (1|year), 
                family = binomial, data = ha_cf)
```

## Random effects

```{r}
m_surv2 <- glmer(surv ~ log_shts_prev + density_0_10 + density_10_20 + (1|plot) + (1|year), 
                family = binomial, data = ha_cf)
AICtab(m_surv, m_surv2)
```
Can remove plant ID
```{r}
m_surv3 <- glmer(surv ~ log_shts_prev + density_0_10 + density_10_20 + (1|year), 
                family = binomial, data = ha_cf)
AICtab(m_surv2, m_surv3)
```
Can't remove plot

```{r}
m_surv4 <- glmer(surv ~ log_shts_prev + density_0_10 + density_10_20 + (1|plot), 
                family = binomial, data = ha_cf)
AICtab(m_surv2, m_surv4)
```
Can't remove year.

## Density dependence?

```{r}
Anova(m_surv2)
```

Yes, but only immediate grid cell

@@

- try negative binomial distribution for shoot number
- ~~Density should be total shoot number~~
- ~~Should actually be density in previous year~~
- ~~Shoots should be poisson distribution, not log-transformed?~~
- incorporate 1-ha fragments and use habitat type as a predictor.
- Implement FLM approach to density dependence.
